MATLAB: Read Me

### 2020-03-30 MATLAB_Files/MATLAB_readMe/MATLAB_readMe.tex
Added most Lamperti MATLAB files. The parameters are 10 and 0.1, now I need to check that all functions are working as they should.

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Normalization:}

@@

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{SDE first moment (1/2):} \label{m1}

Given some measurement $v_{t_{n-1}}$, we want to compute the first moment at time $t_n$. The exact first moment $m_1(s)$ for $s\in[t_{n-1},t_n]$ is the solution of the ODE

@@

Given some measurement $v_{t_{n-1}}$, we want to compute the second moment at time $t_n$. The exact second moment $m_2(s)$ for $s\in[t_{n-1},t_n]$ is the solution of the ODE

@@

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{From $\theta_0$ to $\theta_t$:}
{\small
To ensure that the analytical solutions is always in $]0,1[$, we choose the drift parameter to be
\begin{equation*}
\theta(t) = \max\left(\theta_0,\frac{\theta_0\alpha+2|\dot{p}^\epsilon(t)|}{2\min(p^\epsilon(t),1-p^\epsilon(t))}\right),\quad\theta_0>0.
\end{equation*}
\quad\\
\quad\\
\begin{center}
\begin{tabular}{|c|}
\toprule
{\tiny
\lstinputlisting{../theta_t.m}
}\\
\bottomrule
\end{tabular}
\end{center}}

@@

\alert{This function is independent of if we are using the real data or the truncated one.} If we take a total of $Z\in\N$ days, the batch corresponding to this days is
\begin{table}[]
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
\multicolumn{7}{|l|}{PATH 1}                                                                                       & ...                  & PATH Z               \\ \hline
\multicolumn{2}{|l|}{$t_n=01:10$}   & \multicolumn{2}{l|}{$t_n=01:20$}    & ... & \multicolumn{2}{l|}{$t_n=00:50$} & \multirow{4}{*}{...} & \multirow{4}{*}{...} \\ \cline{1-7}
$p(t_{n-1})$      & $p(t_{n})$      & $p(t_{n-1})$      & $p(t_{n})$      & ... & $p(t_{n-1})$     & $p(t_{n})$    &                      &                      \\ \cline{1-7}
$\dot{p}(t_{n-1})$ & $\dot{p}(t_{n})$ & $\dot{p}(t_{n-1})$ & $\dot{p}(t_{n})$ & ... & $\dot{p}(t_{n-1})$     & $\dot{p}(t_{n})$    &                      &                      \\ \cline{1-7}

@@

>>>>

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Normalization:}\label{Norm}

@@

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{SDE first moment (1/2):} \label{m1}
\alert{Check if still correct for the new model!!!}\\
Given some measurement $v_{t_{n-1}}$, we want to compute the first moment at time $t_n$. The exact first moment $m_1(s)$ for $s\in[t_{n-1},t_n]$ is the solution of the ODE

@@

\alert{Check if still correct for the new model!!!}\\
Given some measurement $v_{t_{n-1}}$, we want to compute the second moment at time $t_n$. The exact second moment $m_2(s)$ for $s\in[t_{n-1},t_n]$ is the solution of the ODE

@@

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{From $\theta_0$ to $\theta_t$:}

To ensure that the analytical solutions is always in $]0,1[$, we choose the drift parameter to be
\begin{equation*}
\theta(t) = \max\left(\theta_0,\frac{\theta_0\alpha+2|\dot{p}^\epsilon(t)|}{2\min(p^\epsilon(t),1-p^\epsilon(t))}\right),\quad\theta_0>0.
\end{equation*}
\quad\\
\quad\\
\begin{center}
\begin{tabular}{|c|}
\toprule
{\tiny
\lstinputlisting{../theta_t.m}
}\\
\bottomrule
\end{tabular}
\end{center}
{\tiny Recall that we normalized w.r.t. time $T$. The normalized $\hat{\theta}_t$ is $\hat{\theta}_t=\theta_tT$. Notice that in the definition of $\theta_t$, if we use $\hat{\theta}_0=\theta_0T$ and the derivative of $p_t$ w.r.t. $\dif \hat{t}$ instead of w.r.t. $\dif t$, then we automatically have $\hat{\theta}_t$. See slide \ref{Norm}.}

@@

\alert{This function is independent of if we are using the real data or the truncated one.}\\
If we take a total of $Z\in\N$ days, the batch corresponding to this days is
\begin{table}[]
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
\multicolumn{7}{|l|}{PATH 1}                                                                                       & ...                  & PATH Z               \\ \hline
\multicolumn{2}{|l|}{$t_n=01:10$}   & \multicolumn{2}{l|}{$t_n=01:20$}    & ... & \multicolumn{2}{l|}{$t_n=00:50$} & \multirow{4}{*}{...} & \multirow{4}{*}{...} \\ \cline{1-7}
$p(t_{n-1})$      & $p(t_{n})$      & $p(t_{n-1})$      & $p(t_{n})$      & ... & $p(t_{n-1})$     & $p(t_{n})$    &                      &                      \\ \cline{1-7}
$\dot{p}(t_{n-1})$ & $\dot{p}(t_{n})$ & $\dot{p}(t_{n-1})$ & $\dot{p}(t_{n})$ & ... & $\dot{p}(t_{n-1})$     & $\dot{p}(t_{n})$    &                      &                      \\ \cline{1-7}

@@

\setbeamercolor{background canvas}{bg=green!20}
\begin{frame}

{\Huge Lamperti Transform}

\setbeamercolor{background canvas}{bg=white!20}
\begin{frame}\frametitle{Two candidates:}

\begin{columns}

\column{.5\textwidth}
Recall $X_t=V_t+p_t$.\\
\quad\\
We have the candidates:
\begin{itemize}

\item {\color{blue} $Z_t=\frac{1}{\sqrt{2\alpha\theta_0}}\arcsin(2(V_t+p_t)-1)$.}
\item {\color{red} $Z_t=\sqrt{\frac{2}{\alpha\theta_0}}\arcsin(\sqrt{1-V_t-p_t})$.}

\end{itemize}
\quad\\
\quad\\
They have the same partial derivatives. The main difference appears in the resulting SDE for $Z_t$. In the {\color{blue} first case}, we have a $\sin(\sqrt{2\alpha\theta_0}Z_t)$ in the drift, while in the \alert{second case}, we have $\cos(\sqrt{2\alpha\theta_0}Z_t)$.\\
\quad\\
For now, we will use the {\color{blue}first candidate}.

\column{.5\textwidth}
\begin{figure}[ht!]
\centering
\includegraphics[width=0.9\textwidth]{../../Mathematica_Files/new_model/Lamp_Comp.pdf}
\end{figure}

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Lamperti data:}

We are following the {\color{blue}first candidate}. Then, we caa

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Lamperti moments:}

The Lamperti SDE is:
\begin{align*}
\dif Z_t= \underbrace{\left[  \frac{(\alpha \theta_0 - \theta_t) {\color{blue}\sin(\sqrt{2 \alpha \theta_0 } Z_t) }- \theta_t (1 - 2 p_t) + 2  \dot{p}_t }{\sqrt{2 \alpha \theta_0} \cos{(\sqrt{2 \alpha \theta_0} Z_t)}}  \right]}_{\coloneqq b(Z_t)} \dif t + \dif W_t . 
\end{align*}
Here we follow Bayesian Filtering and Smoothing, Chapter 9. We use the linearization-based approximation for the moments.

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{SDE first Lamperti moment (mean) (1/2):}
Given some measurement $z_{t_{n-1}}$, we want to compute the first moment at time $t_n$. The linearly approximated first moment $\mu_Z(s)$ for $s\in[t_{n-1},t_n]$ is the solution of the ODE

\begin{equation*}
\begin{cases}
\dif \mu_L(s)&=b(\mu_L(s))\dif s,\\
\mu_L(t_{n-1})&=z_{t_{n-1}}.
\end{cases}
\end{equation*}
\quad\\
We solve numerically the ODE using Forward-Euler:
\begin{equation*}
\mu_L(s_{n})=\mu_L(s_{n-1})+b(\mu_L(s_{n-1}))\Delta s.
\end{equation*}

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{SDE first Lamperti moment (mean) (2/2):}

\begin{center}
\begin{tabular}{|c|}
\toprule
{\scriptsize
\lstinputlisting{../moment_1_L.m}
}\\
\bottomrule
\end{tabular}
\end{center}

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{SDE Lamperti variance (1/2):}
Given some measurement $z_{t_{n-1}}$, we want to compute the variance at time $t_n$. The linearly approximated variance $v_Z(s)$ for $s\in[t_{n-1},t_n]$ is the solution of the ODE

\begin{equation*}
\begin{cases}
\dif \sigma^2_L(s)&=\left(2\sigma^2_Lb'(\mu_L(s))+1\right)\dif s,\\
\sigma^2_L(t_{n-1})&=0.
\end{cases}
\end{equation*}
\quad\\
We solve numerically the ODE using Forward-Euler:
\begin{equation*}
\sigma^2_L(s_{n})=\sigma^2_L(s_{n-1})+\left(1+2\sigma^2_L(s_{n-1})b'(\mu_L(s_{n-1}))\right)\Delta s.
\end{equation*}
We have that (computed with Mathematica):
\begin{equation*}
b'(z)=\frac{\dif b}{\dif z}(z)=\frac{\alpha\theta_0-\theta_t+(2\dot{p}_t+(2p_t-1)\theta_t)\sin(z\sqrt{2\alpha\theta_0}))}{\cos^2(z\sqrt{2\alpha\theta_0})}.
\end{equation*}

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{SDE Lamperti variance (2/2):}

\begin{center}
\begin{tabular}{|c|}
\toprule
{\tiny
\lstinputlisting{../moment_2_L.m}
}\\
\bottomrule
\end{tabular}
\end{center}

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Density next measurement:}

We want the next measurement $Z_{t_n}|Z_{t_{n-1}}$ to have a Gaussian with mean $\mu_Z$ and variance $\sigma_Z^2$ ($f_Z(z)=\mathcal{N}(\mu_Z,\sigma^2_Z)$).\\
\quad\\
Then, we want the SDE and our new PDF $f_Z(z)$ to have the same moments at each $t\in\{\text{some appropriate domain}\}$, i.e., $\mu_Z(t)=\mu_L(t)$ and $\sigma^2_Z(t)=\sigma^2_L(t)$.\\
\quad\\
The density and log-density (natural logarithm) are:
\begin{equation*}
f_Z(z)=\frac{e^{-\frac{1}{2}\left(\frac{z-\mu_Z}{\sigma_Z}\right)^2}}{\sigma_Z\sqrt{2\pi}}\iff \log(f_Z(z))=-\frac{1}{2}\left(\frac{z-\mu_Z}{\sigma_Z}\right)^2-\log(\sigma_Z\sqrt{2\pi}).
\end{equation*}

\begin{center}
\begin{tabular}{|c|}
\toprule
{\tiny
\lstinputlisting{../log_dist_L.m}
}\\
\bottomrule
\end{tabular}
\end{center}
